# Figure 1
```{r}
# Create a QC plot after filtering, using the same SpatialTrans_DMSO_ONTgene object
SpatPlot2_Leiden_clusters <- spatPlot2D(
  gobject = SpatialTrans_DMSO_ONTgene,  # Use the same object with statistics
  cell_color = "leiden_clus",           # Color based on the number of features (genes) per spot
  point_size = 4.5,
  point_alpha = 0.5,
  show_image = TRUE,                 # Show the underlying image
  save_plot = TRUE,                  # Save the plot
  title = "",                        # Remove the title
  save_param = base_save_parameters,   # Apply the save parameters
  return_plot = TRUE,                 # Return the plot object to display after saving
  default_save_name = "SpatPlot2_Leiden_clusters"  # Save plot with this name
)

```

```{r}
# Create the heatmap
heatmap_scranMarkers <- plotMetaDataHeatmap(
  SpatialTrans_DMSO_ONTgene, 
  selected_feats = unique(topgenes_scran),
  metadata_cols = "leiden_clus",
  custom_cluster_order = custom_order,
  x_text_size = 22, 
  y_text_size = 22,
  save_param = modifyList(base_save_parameters, list(base_width = 4, base_height = 12)),
  return_plot = TRUE
)

# Remove y-axis title, change x-axis title, and change legend title
heatmap_scranMarkers <- heatmap_scranMarkers +
  labs(y = NULL, x = "Leiden Clusters", fill = "Leiden Clusters")

# Display the heatmap
print(heatmap_scranMarkers)




```


EnrichR analysis
```{r}
# Required Libraries
library(ggplot2)
library(cowplot)
library(stringr)  # To wrap text
library(enrichR)  # Assuming enrichR is used for enrichment analysis

# Get top 30 significant genes per cluster
top_genes_per_cluster <- scran_markers[, head(.SD, 200), by = cluster, .SDcols = "feats"]

# Define Enrichr libraries of interest
enrichr_libraries <- c("GO_Biological_Process_2023")

# Initialize a list to store enrichment results
enrichment_results <- list()

# Loop over each cluster and perform enrichment
for (cluster_id in unique(top_genes_per_cluster$cluster)) {
  
  # Extract the top 30 gene list for the current cluster
  gene_list <- top_genes_per_cluster[cluster == cluster_id, feats]
  
  # Perform enrichment analysis using Enrichr
  result <- enrichr(genes = gene_list, databases = enrichr_libraries)
  
  # Store the result for the cluster
  enrichment_results[[cluster_id]] <- result
  
  # Print a message to indicate completion of each cluster
  cat("Completed enrichment analysis for cluster:", cluster_id, "\n")
}

```

Plot the enrichment plots


```{r}
#Modified Gene Enrichment Plot Function

# Function to customize the plot with larger white text and correct color scaling and bar order
# Modified Gene Enrichment Plot Function with Larger Legend
custom_plotEnrich <- function(cluster_result, cluster_id) {
  
  # Extract count from Overlap column (assuming format like "5/200")
  cluster_result$Count <- as.numeric(sapply(strsplit(cluster_result$Overlap, "/"), `[`, 1))
  
  # Take top 5 terms ordered by P.value
  cluster_result <- head(cluster_result[order(cluster_result$P.value), ], 5)
  
  # Calculate the maximum bar width for proper spacing
  max_count <- max(cluster_result$Count, na.rm = TRUE)
  
  # Create the plot
  p <- ggplot(cluster_result, aes(x = Count, y = reorder(Term, -P.value))) +  # Reverse order of terms by P.value
    # Add invisible bars to reserve space
    geom_bar(stat = "identity", fill = "transparent", width = 0.7) +
    # Add colored bars, using P.value directly for the fill
    geom_bar(stat = "identity", aes(fill = P.value), width = 0.9) +
    # Add text labels at a consistent position with increased size and white color
    geom_text(aes(x = 0.1, label = Term), 
              hjust = 0,
              size = 20 / .pt,  # Convert to points for actual size 20
              color = "white",
              fontface = "bold",  # Make text bold
              position = position_nudge(y = 0)) +
    # Set proper plot limits
    scale_x_continuous(limits = c(0, max_count * 1.2), 
                       expand = c(0, 0)) +
    # Color scale for P-values: smaller p-values as red, larger as blue
    scale_fill_gradient(low = "red", high = "blue", guide = "colorbar") + 
    labs(title = paste("Cluster", cluster_id),
         x = "Gene Count",
         fill = "P-value") +  # Update legend label to "P-value"
    theme_minimal() +
    theme(
      axis.text.y = element_blank(),
      axis.title.y = element_blank(),
      panel.grid = element_blank(),
      plot.title = element_text(size = 17, face = "bold"),
      axis.text.x = element_text(size = 8),
      axis.title.x = element_text(size = 12),
      legend.position = "right",
      legend.key.size = unit(2, "lines"),  # Increase legend key size
      legend.text = element_text(size = 10),  # Increase legend text size
      legend.title = element_text(size = 12)  # Increase legend title size
    )
  
  return(p)
}

# Create plots with error handling
plots <- list()
for (cluster_id in names(enrichment_results)) {
  cluster_result <- enrichment_results[[cluster_id]][["GO_Biological_Process_2023"]]
  plot <- tryCatch({
    custom_plotEnrich(cluster_result, cluster_id)
  }, error = function(e) {
    message(sprintf("Error in cluster %s: %s", cluster_id, e$message))
    return(NULL)
  })
  
  if (!is.null(plot)) {
    plots[[cluster_id]] <- plot
  }
}

# Remove NULL entries
plots <- plots[!sapply(plots, is.null)]

# Only create combined plot if we have plots to combine
# Create combined plot if we have plots to combine
if (length(plots) > 0) {
  # Combine plots
  GOBiologicalProcess_combined <- plot_grid(
    plotlist = plots, 
    ncol = 1,
    align = 'v',
    axis = 'l'
  )
  
  # Add the title to the combined plot
  GOBiologicalProcess_combined <- plot_grid(
    ggdraw() + 
      draw_label(
        "Top 5 GO Biological Process Enrichment", 
        fontface = 'bold', 
        size = 30,   # Adjust the size of the title
        hjust = 0.5  # Center the title horizontally
      ),
    GOBiologicalProcess_combined,
    ncol = 1, 
    rel_heights = c(0.1, 1)  # Adjust the relative height of the title
  )

  # Save the plot with the title
  ggsave(
    filename = "./figures/GOBiologicalProcess_combined.png",
    plot = GOBiologicalProcess_combined,
    width = 15, 
    height = 5 * length(plots) + 1,  # Adjust height based on number of plots and space for title
    dpi = 600
  )
  
  # Display the combined plot with title (optional)
  print(GOBiologicalProcess_combined)
} else {
  message("No valid plots were created. Please check your enrichment results.")
}
# Display the adjusted plot (optional)
print(GOBiologicalProcess_combined)

```

#  Save the enrichment result analysis 

```{r}
# Initialize an empty list to store flattened results
flattened_results <- list()

# Loop over each cluster and flatten the enrichment results
for (cluster_id in names(enrichment_results)) {
  if (!is.null(enrichment_results[[cluster_id]][["GO_Biological_Process_2023"]])) {
    cluster_data <- enrichment_results[[cluster_id]][["GO_Biological_Process_2023"]]
    cluster_data$Cluster <- cluster_id  # Add cluster ID as a new column
    
    # Concatenate gene names into a single string
    cluster_data$Genes <- sapply(cluster_data$Genes, function(genes) {
      paste(genes, collapse = ", ")
    })
    
    flattened_results[[cluster_id]] <- cluster_data
  }
}

# Combine all cluster results into a single data frame
combined_results <- do.call(rbind, flattened_results)

# Define the file path for saving
enrichment_results_file <- file.path(supplementary_figures, "Supplementary Table S2.csv")

# Save the combined results as a CSV file
fwrite(
  combined_results,
  file = enrichment_results_file
)

# Confirmation message
message("Enrichment results saved to: ", enrichment_results_file)


```

```{r}
# Create the final combined figure
combined_Figure1 <- plot_grid(
  combined_tumoroidPlot, 
  heatmap_scranMarkers,
  GOBiologicalProcess_combined,  # Updated GO Biological Process plot with white background
  labels = c("A", "B", "C"),     # Label plots as A, B, C
  label_size = 30,               # Increase label size for better visibility
  ncol = 3,                      # Arrange plots in three columns
  rel_widths = c(0.8, 0.3, 0.5)    # Adjust relative widths for better balance
)

# Save the combined plot as a PNG file with 600 dpi resolution
ggsave(
  filename = "./figures/paper_figures/Figure_1.png",  # Save plot with a unique name
  plot = combined_Figure1,
  width = 35,                  # Width of the figure in inches
  height = 22,                  # Height of the figure in inches
  dpi = 600                    # Set the resolution to 600 dpi
)

# Display the combined plot (optional)
print(combined_Figure1)


```


# Figure 2
```{r}
# Extract top 5 genes by ranking for cluster 2
top_5_cluster_2 <- scran_markers[cluster == "2"][order(Top)][1:5]

# Extract top 5 genes by ranking for cluster 3
top_5_cluster_3 <- scran_markers[cluster == "3"][order(Top)][1:5]

# Combine the top 5 genes from both clusters
topgenes_scran_5_clus2_3 <- rbind(top_5_cluster_2, top_5_cluster_3)$feats

# Extract the top 8 genes for cluster 4
#Spatial expression patterns of the top genes changed in the clusters

# Define the plot object
spatial_topscran_genes <- spatFeatPlot2D(SpatialTrans_DMSO_ONTgene, 
               expression_values = "scaled",
               feats = topgenes_scran_5_clus2_3,
               show_image = TRUE, 
               cow_n_col = 5, 
               point_size = 2.5,
               point_alpha = 0.7,
               save_param = modifyList(base_save_parameters, list(base_width = 20, base_height = 10)),    # Override base width and height
               return_plot = TRUE,               # Return the plot object to display after saving
               save_plot = FALSE,                # Disable auto-save in the function
               default_save_name = "spatial_topscran_genes"  # Save plot with this name
)

# Save the plot as a PNG with 600 DPI
ggsave(
  filename = "./figures/figure_2.png",     # Path and filename
  plot = spatial_topscran_genes,           # Plot object
  width = 20, height = 10,                 # Width and height
  dpi = 600                                # Set DPI to 600
)

# Display the plot
print(spatial_topscran_genes)

```

# Plotting using `spatFeatPlot2D()` Cannot add scale bar or probaarly crop

```{r}

SpatialTrans_DMSO_ONTgene_subset <- subsetGiottoLocs(SpatialTrans_DMSO_ONTgene,
x_min = 12900,
x_max = 22400,
y_min =-17200,
y_max =-14000
)

# Define the custom list of genes using c()
custom_genes <- c("TTTY14", "MT-ND4",
                  "MUC1", "MT-ND3",
                  "TACSTD2", "MT-ATP6",
                  "PDZK1IP1","MT-ATP8",
                   "MMP7", "COX7B")

# Define the plot object
spatial_topscran_genes <- spatFeatPlot2D(SpatialTrans_DMSO_ONTgene_subset, 
               expression_values = "scaled",
               feats = custom_genes,
               show_image = TRUE, 
               cow_n_col = 2, 
               point_size = 4.5,
               point_alpha = 0.7,
               save_param = modifyList(base_save_parameters, list(base_width = 20, base_height = 10)),    # Override base width and height
               return_plot = TRUE,               # Return the plot object to display after saving
               save_plot = FALSE
)

# Save the plot as a PNG with 600 DPI
ggsave(
  filename = "./figures/paper_figures/figure_2.png",     # Path and filename
  plot = spatial_topscran_genes,           # Plot object
  width = 10, height = 11,                 # Width and height
  dpi = 600                                # Set DPI to 600
)

# Display the plot
print(spatial_topscran_genes)

```

```{r}
# Define the custom list of genes using c()
custom_genes <- c("TTTY14", "MUC1", "TACSTD2", "MMP7", "PDZK1IP1",
                  "MT-ND4", "MT-ND3", "MT-ATP6","MT-ATP8","COX7B")

# Define the plot object
spatial_topscran_genes <- spatFeatPlot2D(SpatialTrans_DMSO_ONTgene, 
               expression_values = "scaled",
               feats = custom_genes,
               show_image = TRUE, 
               cow_n_col = 5, 
               point_size = 2.5,
               point_alpha = 0.7,
               save_param = modifyList(base_save_parameters, list(base_width = 20, base_height = 10)),    # Override base width and height
               return_plot = TRUE,               # Return the plot object to display after saving
               save_plot = FALSE
)

# Save the plot as a PNG with 600 DPI
ggsave(
  filename = "./figures/paper_figures/supplementary/figure_2_wholesections.png",     # Path and filename
  plot = spatial_topscran_genes,           # Plot object
  width = 23, height = 11,                 # Width and height
  dpi = 600                                # Set DPI to 600
)

# Display the plot
print(spatial_topscran_genes)
```

# Subset and plots using modified `spatInSituPlotPoints`
```{r}
spatinsitu_feat_workaround <- function(
        gobject, feat, 
        expression_values = "scaled", 
        spat_unit = NULL, 
        feat_type = NULL, 
        ...) {
    expr_table <- spatValues(gobject,
                             spat_unit = spat_unit, 
                             feat_type = feat_type,
                             expression_values = expression_values,
                             feats = feat)
    gobject <- addCellMetadata(gobject,
                               spat_unit = spat_unit,
                               feat_type = feat_type,
                               new_metadata = expr_table,
                               by_column = TRUE)
    p <- list(
        gobject = gobject,
        polygon_fill = feat,
        ...
    )
    do.call(spatInSituPlotPoints, p)
}
```


```{r}
# Required Libraries
library(ggplot2)
library(patchwork)

# Define the list of genes
genes <- c("TTTY14", "MT-ND4", "MUC1", "MT-ND3", "TACSTD2", "MT-ATP6", "PDZK1IP1", "MT-ATP8", "MMP7", "COX7B")

# Define the fixed scale bar length in microns
scale_bar_length_microns <- 100  
scale_bar_length_pixels <- scale_bar_length_microns / microns_per_pixel

# Coordinates for placing the scale bar
x_right <- 22000 - scale_bar_length_pixels - 20  # Adjusted for xlim
y_bottom <- -17200 + 80  # Adjusted for ylim

# List to store plots
gene_plots <- list()

# Generate plots for each gene
for (gene in genes) {
  # Generate the plot for the current gene
  gene_plot <- spatinsitu_feat_workaround(
      SpatialTrans_DMSO_ONTgene, 
      feat =  c(gene),
      show_image = TRUE, 
      image_name = "image",
      return_plot = TRUE,
      save_param = modifyList(base_save_parameters, list(base_width = 20, base_height = 10)),
      save_plot = FALSE,
      polygon_alpha = 0.7,
      polygon_line_size = 0.1,
      polygon_color = "#333030",
      background_color = "white", 
      xlim = c(13600, 22000), 
      ylim = c(-17200, -13800)
  ) +
    ggplot2::theme(
      axis.line = ggplot2::element_line(color = "black"),
      axis.title.x = ggplot2::element_text(size = 10),
      axis.title.y = ggplot2::element_text(size = 10),
      plot.title = ggplot2::element_text(size = 12, hjust = 0.5)  # Title formatting
    ) +
    ggplot2::labs(
      x = "x coordinates",
      y = "y coordinates",
      title = gene  # Use the current gene name as the title
    ) +
    ggplot2::annotate(
      "segment",
      x = x_right,
      xend = x_right + scale_bar_length_pixels,
      y = y_bottom + 110,
      yend = y_bottom + 110,
      color = "black",
      linewidth = 0.8,
      lineend = "round"
    ) +
    ggplot2::annotate(
      "text",
      x = x_right + scale_bar_length_pixels / 2,
      y = y_bottom - 20,
      label = paste0(scale_bar_length_microns, " µm"),
      color = "black",
      size = 2,
      hjust = 0.5
    )
  
  # Save the plot to the list
  gene_plots[[gene]] <- gene_plot
  
  # Save the individual plot as an image
  ggsave(
    filename = paste0("./results/Giotto_DMSO_results/ONTgene/", gene, "_spatial_expression.png"),
    plot = gene_plot,
    width = 20,
    height = 10,
    dpi = 600
  )
}

# Split the plots into two columns
column1_genes <- c("TTTY14", "MUC1", "TACSTD2", "PDZK1IP1", "MMP7")
column2_genes <- c("MT-ND4", "MT-ND3", "MT-ATP6", "MT-ATP8", "COX7B")

# Combine plots into two columns
final_grid <- (patchwork::wrap_plots(gene_plots[column1_genes], ncol = 1) | 
               patchwork::wrap_plots(gene_plots[column2_genes], ncol = 1))

# Save the combined grid as an image
combined_save_path <- "./figures/paper_figures/figure_2.png"
ggsave(
  filename = combined_save_path,
  plot = final_grid,
  width = 10,  # Adjust dimensions as needed
  height = 11, 
  dpi = 600
)

# Print the combined grid
print(final_grid)

```

# Plot the differntial expression genes Control vs NUC7738
## Plot DMSO part

```{r}
# Required Libraries
library(ggplot2)
library(patchwork)

# Define the list of genes
genes <- c("SNHG25", "HLA-B", "B2M", "MT-CO1", "RPL37", "POLR2L")

# Define the fixed scale bar length in microns
scale_bar_length_microns <- 100  
scale_bar_length_pixels <- scale_bar_length_microns / microns_per_pixel

# Initialize the gene_plots list
gene_plots <- list()

# Plot DMSO part
x_right <- 22000 - scale_bar_length_pixels - 20  # Adjusted for xlim
y_bottom <- -17200 + 80  # Adjusted for ylim
for (gene in genes) {
  dmso_plot <- spatinsitu_feat_workaround(
    SpatialTrans_combined_ONTgene, 
    feat = c(gene),
    show_image = TRUE, 
    show_legend = FALSE,
    image_name = c("DMSO-image"),
    return_plot = TRUE,
    save_param = modifyList(base_save_parameters, list(base_width = 20, base_height = 10)),
    save_plot = FALSE,
    polygon_alpha = 0.8,
    polygon_line_size = 0.1,
    polygon_color = "#333030",
    background_color = "white", 
    xlim = c(13600, 22000), 
    ylim = c(-17200, -13800)
  ) +
    ggplot2::theme(
      axis.line = ggplot2::element_line(color = "black"),
      axis.title.x = ggplot2::element_text(size = 10),
      axis.title.y = ggplot2::element_text(size = 10),
      plot.title = ggplot2::element_text(size = 12, hjust = 0.5),  # Title formatting
      legend.position = "none"  # Explicitly remove the legend
    ) +
    ggplot2::labs(
      x = "x coordinates",
      y = "y coordinates"
    ) +
    ggplot2::annotate(
      "segment",
      x = x_right,
      xend = x_right + scale_bar_length_pixels,
      y = y_bottom + 110,
      yend = y_bottom + 110,
      color = "black",
      linewidth = 0.8,
      lineend = "round"
    ) +
    ggplot2::annotate(
      "text",
      x = x_right + scale_bar_length_pixels / 2,
      y = y_bottom - 20,
      label = paste0(scale_bar_length_microns, " µm"),
      color = "black",
      size = 2,
      hjust = 0.5
    )
  
  # Add to the gene_plots list
  gene_plots[[paste0("DMSO_", gene)]] <- dmso_plot

  # Save the individual plot
  ggsave(
    filename = paste0("./results/SpatialTrans_combined_ONTgene/ONTgene/", gene, "_DMSO_spatial_expression.png"),
    plot = dmso_plot,  # Corrected variable name
    width = 20,
    height = 10,
    dpi = 600
  )
}

```


## Plot NUC7738 part

```{r}
# Plot NUC7738 part
x_right <- 49900 - scale_bar_length_pixels - 20  # Adjusted for xlim
y_bottom <- -25200 + 80  # Adjusted for ylim
for (gene in genes) {
  nuc7738_plot <- spatinsitu_feat_workaround(
    SpatialTrans_combined_ONTgene, 
    feat = c(gene),
    show_image = TRUE, 
    image_name = c("30uM7738-image"),
    return_plot = TRUE,
    save_param = modifyList(base_save_parameters, list(base_width = 20, base_height = 10)),
    save_plot = FALSE,
    polygon_alpha = 0.8,
    polygon_line_size = 0.1,
    polygon_color = "#333030",
    background_color = "white", 
    xlim = c(40900, 49900), 
    ylim = c(-25200, -21500)
  ) +
    ggplot2::theme(
      axis.line = ggplot2::element_line(color = "black"),
      axis.title.x = ggplot2::element_text(size = 10),
      axis.title.y = ggplot2::element_text(size = 10),
      plot.title = ggplot2::element_text(size = 12, hjust = 0.5)  # Title formatting
    ) +
    ggplot2::labs(
      x = "x coordinates",
      y = "y coordinates"
    ) +
    ggplot2::annotate(
      "segment",
      x = x_right,
      xend = x_right + scale_bar_length_pixels,
      y = y_bottom + 110,
      yend = y_bottom + 110,
      color = "black",
      linewidth = 0.8,
      lineend = "round"
    ) +
    ggplot2::annotate(
      "text",
      x = x_right + scale_bar_length_pixels / 2,
      y = y_bottom - 20,
      label = paste0(scale_bar_length_microns, " µm"),
      color = "black",
      size = 2,
      hjust = 0.5
    )
  gene_plots[[paste0("NUC7738_", gene)]] <- nuc7738_plot
  ggsave(
   filename = paste0("./results/SpatialTrans_combined_ONTgene/ONTgene/", gene, "_NUC7738_spatial_expression.png"),
   plot = gene_plot,
   width = 20,
   height = 10,
   dpi = 600
  )
}

```


```{r}
library(ggplot2)
library(patchwork)

# Combine DMSO and NUC7738 plots into a grid
combined_plots <- lapply(genes, function(gene) {
  dmso_plot <- gene_plots[[paste0("DMSO_", gene)]]
  nuc7738_plot <- gene_plots[[paste0("NUC7738_", gene)]]
  dmso_plot | nuc7738_plot
})

# Create gene labels (one for each row)
gene_labels <- lapply(genes, function(gene) {
  ggplot() +
    ggplot2::annotate("text", x = 0.5, y = 0.5, label = gene, size = 6, fontface = "bold", angle = 90) +
    theme_void() +
    theme(plot.margin = margin(0, 0, 0, 0))  # Minimize margin
})

# Combine gene labels and their respective plots row by row
labeled_rows <- mapply(
  function(label, plot) {
    wrap_plots(label, plot, widths = c(0.03, 1))  # Assign a smaller width for labels
  },
  gene_labels,
  combined_plots,
  SIMPLIFY = FALSE
)

# Combine all rows into a single column layout
labeled_grid <- wrap_plots(labeled_rows, ncol = 1)

# Create column labels for DMSO and NUC7738
column_labels <- wrap_plots(
  ggplot() +
    ggplot2::annotate("text", x = 0.5, y = 0.5, label = "Control (DMSO)", size = 8, fontface = "bold") +
    theme_void() +
    theme(plot.margin = margin(0, 0, 0, 0)),  # Minimize margin
  ggplot() +
    ggplot2::annotate("text", x = 0.5, y = 0.5, label = "NUC-7738", size = 8, fontface = "bold") +
    theme_void() +
    theme(plot.margin = margin(0, 0, 0, 0)),  # Minimize margin
  ncol = 2
)

# Combine column labels with the plot grid
final_grid <- column_labels / labeled_grid +
  plot_layout(heights = c(0.02, 1))  # Adjust height for column labels

# Save the final grid with column and row labels
combined_save_path <- "./figures/paper_figures/figure_4_6genes.png"
ggsave(
  filename = combined_save_path,
  plot = final_grid,
  width = 12,  # Adjust width to balance labels and plots
  height = 16, # Adjust height to fit all gene pairs
  dpi = 600
)

# Print the final grid
print(final_grid)
```



# Plot whole section

# Plot the differntial expression genes Control vs NUC7738
## Plot DMSO part

```{r}
# Required Libraries
library(ggplot2)
library(patchwork)

# Define the list of genes
genes <- c("MT-CO1", "RPL37", "POLR2L")

# Define the fixed scale bar length in microns
scale_bar_length_microns <- 250  
scale_bar_length_pixels <- scale_bar_length_microns / microns_per_pixel

# Initialize the gene_plots list
gene_plots <- list()

# Plot DMSO part
x_right <- 23000 - scale_bar_length_pixels - 20  # Adjusted for xlim
y_bottom <- -22200 + 80  # Adjusted for ylim
for (gene in genes) {
  dmso_plot <- spatinsitu_feat_workaround(
    SpatialTrans_combined_ONTgene, 
    feat = c(gene),
    show_image = TRUE, 
    show_legend = FALSE,
    image_name = c("DMSO-image"),
    return_plot = TRUE,
    save_param = modifyList(base_save_parameters, list(base_width = 20, base_height = 10)),
    save_plot = FALSE,
    polygon_alpha = 0.8,
    polygon_line_size = 0.1,
    polygon_color = "#333030",
    background_color = "white", 
    xlim = c(13800, 23000), 
    ylim = c(-22200, -8500)
  ) +
    ggplot2::theme(
      axis.line = ggplot2::element_line(color = "black"),
      axis.title.x = ggplot2::element_text(size = 10),
      axis.title.y = ggplot2::element_text(size = 10),
      plot.title = ggplot2::element_text(size = 12, hjust = 0.5),  # Title formatting
      legend.position = "none"  # Explicitly remove the legend
    ) +
    ggplot2::labs(
      x = "x coordinates",
      y = "y coordinates"
    ) +
    ggplot2::annotate(
      "segment",
      x = x_right,
      xend = x_right + scale_bar_length_pixels,
      y = y_bottom + 350,
      yend = y_bottom + 350,
      color = "black",
      linewidth = 0.5,
      lineend = "round"
    ) +
    ggplot2::annotate(
      "text",
      x = x_right + scale_bar_length_pixels / 2,
      y = y_bottom - 20,
      label = paste0(scale_bar_length_microns, " µm"),
      color = "black",
      size = 2,
      hjust = 0.5
    )
  
  # Add to the gene_plots list
  gene_plots[[paste0("DMSO_", gene)]] <- dmso_plot

  # Save the individual plot
  ggsave(
    filename = paste0("./results/SpatialTrans_combined_ONTgene/ONTgene/", gene, "_DMSO_spatial_expression_whole.png"),
    plot = dmso_plot,  # Corrected variable name
    width = 20,
    height = 10,
    dpi = 600
  )
}

```


## Plot NUC7738 part

```{r}
# Plot NUC7738 part
x_right <- 51900- scale_bar_length_pixels - 20  # Adjusted for xlim
y_bottom <- -26500 + 80  # Adjusted for ylim
for (gene in genes) {
  nuc7738_plot <- spatinsitu_feat_workaround(
    SpatialTrans_combined_ONTgene, 
    feat = c(gene),
    show_image = TRUE, 
    image_name = c("30uM7738-image"),
    return_plot = TRUE,
    save_param = modifyList(base_save_parameters, list(base_width = 20, base_height = 10)),
    save_plot = FALSE,
    polygon_alpha = 0.8,
    polygon_line_size = 0.1,
    polygon_color = "#333030",
    background_color = "white", 
    xlim = c(40900, 51900), 
    ylim = c(-26500, -10500)
  ) +
    ggplot2::theme(
      axis.line = ggplot2::element_line(color = "black"),
      axis.title.x = ggplot2::element_text(size = 10),
      axis.title.y = ggplot2::element_text(size = 10),
      plot.title = ggplot2::element_text(size = 12, hjust = 0.5)  # Title formatting
    ) +
    ggplot2::labs(
      x = "x coordinates",
      y = "y coordinates"
    ) +
    ggplot2::annotate(
      "segment",
      x = x_right,
      xend = x_right + scale_bar_length_pixels,
      y = y_bottom + 350,
      yend = y_bottom + 350,
      color = "black",
      linewidth = 0.5,
      lineend = "round"
    ) +
    ggplot2::annotate(
      "text",
      x = x_right + scale_bar_length_pixels / 2,
      y = y_bottom - 20,
      label = paste0(scale_bar_length_microns, " µm"),
      color = "black",
      size = 2,
      hjust = 0.5
    )
  gene_plots[[paste0("NUC7738_", gene)]] <- nuc7738_plot
  ggsave(
   filename = paste0("./results/SpatialTrans_combined_ONTgene/ONTgene/", gene, "_NUC7738_spatial_whole.png"),
   plot = nuc7738_plot,
   width = 20,
   height = 10,
   dpi = 600
  )
}

```


```{r}
library(ggplot2)
library(patchwork)

# Combine DMSO and NUC7738 plots into a grid
combined_plots <- lapply(genes, function(gene) {
  dmso_plot <- gene_plots[[paste0("DMSO_", gene)]]
  nuc7738_plot <- gene_plots[[paste0("NUC7738_", gene)]]
  dmso_plot | nuc7738_plot
})

# Create gene labels (one for each row)
gene_labels <- lapply(genes, function(gene) {
  ggplot() +
    ggplot2::annotate("text", x = 0.5, y = 0.5, label = gene, size = 6, fontface = "bold", angle = 90) +
    theme_void() +
    theme(plot.margin = margin(0, 0, 0, 0))  # Minimize margin
})

# Combine gene labels and their respective plots row by row
labeled_rows <- mapply(
  function(label, plot) {
    wrap_plots(label, plot, widths = c(0.03, 1))  # Assign a smaller width for labels
  },
  gene_labels,
  combined_plots,
  SIMPLIFY = FALSE
)

# Combine all rows into a single column layout
labeled_grid <- wrap_plots(labeled_rows, ncol = 1)

# Create column labels for DMSO and NUC7738
column_labels <- wrap_plots(
  ggplot() +
    ggplot2::annotate("text", x = 0.5, y = 0.5, label = "Control (DMSO)", size = 8, fontface = "bold") +
    theme_void() +
    theme(plot.margin = margin(0, 0, 0, 0)),  # Minimize margin
  ggplot() +
    ggplot2::annotate("text", x = 0.5, y = 0.5, label = "NUC-7738", size = 8, fontface = "bold") +
    theme_void() +
    theme(plot.margin = margin(0, 0, 0, 0)),  # Minimize margin
  ncol = 2
)

# Combine column labels with the plot grid
final_grid <- column_labels / labeled_grid +
  plot_layout(heights = c(0.02, 1))  # Adjust height for column labels

# Save the final grid with column and row labels
combined_save_path <- "./figures/paper_figures/figure_4_WholseSection_MT-CO1_RPL37_POLR2L.png"
ggsave(
  filename = combined_save_path,
  plot = final_grid,
  width = 10,  # Adjust width to balance labels and plots
  height = 25, # Adjust height to fit all gene pairs
  dpi = 600
)

# Print the final grid
print(final_grid)
```