---
title: "GLS Transcript Isoform Spatial Visualization"
author: "Dr Mustafa Elshani"
date: "`r Sys.Date()`"
output: html_document
---

# GLS Transcript Isoform Spatial Expression

Visualizes GLS (Glutaminase) transcript isoform expression spatially using ONT long-read 
transcript-level data. GLS has two major isoforms: KGA (ENST00000320717) and GAC (ENST00000338435).

---

# 1. Setup

```{r setup, message=FALSE, warning=FALSE}
library(Giotto)
library(ggplot2)
library(cowplot)
library(dplyr)
library(GenomicFeatures)
library(ggtranscript)
library(terra)
library(exifr)
```

# 2. Configuration

```{r config}
# Paths
data_dir_10x_DMSO <- "./DMSO/spaceranger_out_run2/outs"
giotto_object_path <- "./GiottoSuite_DMSO_ONTtrans/gobject.RDS"

# GLS isoforms
GLS_TRANSCRIPTS <- c("ENST00000320717", "ENST00000338435")  # KGA, GAC
GENE_NAME <- "GLS"

# Output
figure_output_dir <- "./figures/paper_figures"
supp_output_dir <- "./figures/paper_figures/supplementary"
```

# 3. Load Data

## 3.1 Load Giotto Object

```{r load-giotto}
SpatialTrans_DMSO_ONTtrans <- loadGiotto(giotto_object_path)
```

## 3.2 Calculate Microns Per Pixel

```{r calculate-resolution}
tiff_file_path <- file.path(data_dir_10x_DMSO, "spatial/tissue_fullres_image.tif")
metadata <- read_exif(tiff_file_path)
microns_per_pixel <- 10000 / metadata$XResolution
print(paste("Microns per pixel:", microns_per_pixel))
```

# 4. Helper Functions

## 4.1 Spatial Feature Plotting Workaround

```{r helper-spatinsitu}
spatinsitu_feat_workaround <- function(
    gobject, feat, 
    expression_values = "scaled", 
    spat_unit = NULL, 
    feat_type = NULL, 
    ...) {
  
  expr_table <- spatValues(gobject,
                           spat_unit = spat_unit, 
                           feat_type = feat_type,
                           expression_values = expression_values,
                           feats = feat)
  
  gobject <- addCellMetadata(gobject,
                             spat_unit = spat_unit,
                             feat_type = feat_type,
                             new_metadata = expr_table,
                             by_column = TRUE)
  
  p <- list(gobject = gobject, polygon_fill = feat, ...)
  do.call(spatInSituPlotPoints, p)
}
```

## 4.2 Spatial Plot Generator

```{r helper-spatial-plot}
create_spatial_plot <- function(giotto_obj, gene, image_name, 
                                 xlim, ylim, x_right, y_bottom,
                                 scale_bar_microns, scale_bar_y_offset = 250,
                                 color_limits = c(0, 2)) {
  
  scale_bar_length_pixels <- scale_bar_microns / microns_per_pixel
  
  p <- spatinsitu_feat_workaround(
    giotto_obj, 
    feat = gene,
    show_image = TRUE, 
    image_name = image_name,
    return_plot = TRUE,
    save_plot = FALSE,
    polygon_alpha = 0.7,
    polygon_line_size = 0.1,
    polygon_color = "#333030",
    background_color = "white", 
    xlim = xlim, 
    ylim = ylim
  ) +
    ggplot2::theme_classic() +
    ggplot2::theme(
      axis.line = ggplot2::element_line(color = "black", linewidth = 1.2),
      axis.title.x = ggplot2::element_text(size = 24),
      axis.title.y = ggplot2::element_text(size = 24),
      axis.text.x = ggplot2::element_text(size = 16),
      axis.text.y = ggplot2::element_text(size = 16, angle = 90, hjust = 0.5),
      plot.title = ggplot2::element_text(size = 26, face = "bold", hjust = 0.5),
      legend.title = ggplot2::element_blank(),
      legend.text = ggplot2::element_text(size = 18),
      legend.key.size = ggplot2::unit(1.2, "cm")
    ) +
    ggplot2::labs(x = "x coordinates", y = "y coordinates", title = gene) +
    ggplot2::scale_fill_gradient(low = "white", high = "red",
                                  limits = color_limits,
                                  oob = scales::squish) +
    ggplot2::annotate("segment",
             x = x_right, xend = x_right + scale_bar_length_pixels,
             y = y_bottom + scale_bar_y_offset, yend = y_bottom + scale_bar_y_offset,
             color = "black", linewidth = 2, lineend = "round") +
    ggplot2::annotate("text",
             x = x_right + scale_bar_length_pixels / 2, y = y_bottom - 20,
             label = paste0(scale_bar_microns, " µm"),
             color = "black", size = 6, fontface = "bold", hjust = 0.5)
  
  return(p)
}
```

## 4.3 Isoform Structure Plot Generator

```{r helper-isoform-plot}
generate_isoform_plot <- function(transcript_ids, txdb) {
  
  isoform_plots <- list()
  transcript_exons <- exonsBy(txdb, by = "tx", use.names = TRUE)
  
  for (tx_id in transcript_ids) {
    specific_exons <- transcript_exons[[which(names(transcript_exons) == tx_id)]]
    
    exon_df <- as.data.frame(specific_exons) %>%
      mutate(transcript_name = "", type = "exon")
    
    isoform_plot <- ggplot(exon_df, aes(xstart = start, xend = end, y = transcript_name)) +
      geom_range(fill = "steelblue", color = "black", alpha = 0.7) +
      geom_intron(
        data = to_intron(exon_df, "transcript_name"),
        aes(strand = strand),
        arrow.min.length = unit(0.05, "npc"),
        arrow = grid::arrow(angle = 30, length = unit(0.1, "inches"))
      ) +
      theme_minimal() +
      theme(
        axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank()
      ) +
      labs(x = "Genomic Coordinates")
    
    isoform_plots[[tx_id]] <- isoform_plot
  }
  
  plot_grid(plotlist = isoform_plots, ncol = length(transcript_ids))
}
```

# 5. Load Ensembl TxDb

```{r load-txdb, message=FALSE, warning=FALSE}
options(ucscChromosomeNames = FALSE)
txdb_ensembl <- makeTxDbFromEnsembl("Homo sapiens", release = 111)
```

# 6. **Figure 3**: GLS Isoform Spatial Expression (Subset Region)

```{r figure3-subset, fig.width=25, fig.height=7}
scale_bar_microns_subset <- 150
scale_bar_pixels_subset <- scale_bar_microns_subset / microns_per_pixel

xlim_subset <- c(13600, 22000)
ylim_subset <- c(-17200, -13800)

# Generate panel labels (A, B, C, D, ...)
panel_labels <- LETTERS[seq_along(GLS_TRANSCRIPTS)]

# Spatial plots for each isoform with panel labels
spatial_plots_subset <- lapply(seq_along(GLS_TRANSCRIPTS), function(i) {
  tx <- GLS_TRANSCRIPTS[i]
  p <- create_spatial_plot(
    giotto_obj = SpatialTrans_DMSO_ONTtrans,
    gene = tx, 
    image_name = "image", 
    xlim = xlim_subset, 
    ylim = ylim_subset, 
    x_right = xlim_subset[2] - scale_bar_pixels_subset - 20, 
    y_bottom = ylim_subset[1] + 80,
    scale_bar_microns = scale_bar_microns_subset,
    scale_bar_y_offset = 250
  )
  return(p)
})

# Combine spatial plots with panel labels
spatial_grid_subset <- plot_grid(
  plotlist = spatial_plots_subset, 
  ncol = length(GLS_TRANSCRIPTS),
  labels = panel_labels,
  label_size = 28,
  label_fontface = "bold"
)

# Isoform structure
isoform_plot_subset <- generate_isoform_plot(GLS_TRANSCRIPTS, txdb_ensembl)

# Combined figure
figure3_plot <- plot_grid(
  spatial_grid_subset,
  isoform_plot_subset,
  ncol = 1,
  rel_heights = c(4, 1)
)

ggsave(
  filename = file.path(figure_output_dir, paste0("figure_3_DMSO_", GENE_NAME, ".png")),
  plot = figure3_plot,
  width = 25, height = 7, dpi = 600, units = "in"
)

ggsave(
  filename = file.path(figure_output_dir, paste0("figure_3_DMSO_", GENE_NAME, ".pdf")),
  plot = figure3_plot,
  width = 25, height = 6.5, units = "in"
)

print(figure3_plot)
```

# 7. Supplementary: GLS Isoform Expression (Whole Section)

```{r supp-whole-section, fig.width=25, fig.height=15}
scale_bar_microns_whole <- 250
scale_bar_pixels_whole <- scale_bar_microns_whole / microns_per_pixel

xlim_whole <- c(13600, 25000)
ylim_whole <- c(-25000, -8500)

# Generate panel labels (A, B, C, D, ...)
panel_labels_whole <- LETTERS[seq_along(GLS_TRANSCRIPTS)]

# Spatial plots for whole section with panel labels
spatial_plots_whole <- lapply(seq_along(GLS_TRANSCRIPTS), function(i) {
  tx <- GLS_TRANSCRIPTS[i]
  p <- create_spatial_plot(
    giotto_obj = SpatialTrans_DMSO_ONTtrans,
    gene = tx, 
    image_name = "image", 
    xlim = xlim_whole, 
    ylim = ylim_whole, 
    x_right = 24000 - scale_bar_pixels_whole - 20, 
    y_bottom = ylim_whole[1] + 80,
    scale_bar_microns = scale_bar_microns_whole,
    scale_bar_y_offset = 400
  )
  return(p)
})

# Combine spatial plots with panel labels
spatial_grid_whole <- plot_grid(
  plotlist = spatial_plots_whole, 
  ncol = length(GLS_TRANSCRIPTS),
  labels = panel_labels_whole,
  label_size = 28,
  label_fontface = "bold"
)

# Isoform structure
isoform_plot_whole <- generate_isoform_plot(GLS_TRANSCRIPTS, txdb_ensembl)

# Combined supplementary figure
supp_plot <- plot_grid(
  spatial_grid_whole,
  isoform_plot_whole,
  ncol = 1,
  rel_heights = c(4, 1)
)

ggsave(
  filename = file.path(supp_output_dir, paste0("figure_3_DMSO_", GENE_NAME, "_wholesection.png")),
  plot = supp_plot,
  width = 25, height = 15, dpi = 600, units = "in"
)

print(supp_plot)
```

# 8. Session Info

```{r session-info}
sessionInfo()
```