
# Load Libraries

```{r}
library(Giotto)
library(dplyr)
library(ggplot2)
library(patchwork)
library(RColorBrewer)
library(terra)
library(exifr)

```

```{r}
# Set the Python path to the existing environment
python_path <- "/home/mustafa/miniforge3/envs/giotto_env/bin/python"

# Set up Giotto instructions with the manually specified Python path
results_folder <- "./results/SpatialTrans_combined_ONTgene/ONTgene"

# Create the results folder if it doesn't exist
if (!dir.exists(results_folder)) {
  dir.create(results_folder, recursive = TRUE)
}

# Create Giotto instructions with Python path
instrs <- createGiottoInstructions(
  save_plot = TRUE, 
  save_dir = results_folder, 
  show_plot = FALSE, 
  python_path = python_path
)

# Check instructions to confirm the setup
print(instrs)

```


# Load Each Visium Section

```{r}
SpatialTrans_DMSO_ONTgene <- loadGiotto("./GiottoSuite_DMSO_ONTgene")
SpatialTrans_30uM7738_ONTgene <- loadGiotto("./GiottoSuite_30uM7738_ONTgene")
```

# Check the load Giotto Objects 
## DMSO
```{r}
# Set up the results folder
SpatialTrans_DMSO_ONTgene_inTissue <- spatPlot2D(gobject = SpatialTrans_DMSO_ONTgene, 
           cell_color = "in_tissue",
           show_image = TRUE, 
           point_size = 2.5,
           cell_color_code = c("0" = "lightgrey", "1" = "blue"),
           point_alpha = 0.5,
           show_plot = TRUE )

print(SpatialTrans_DMSO_ONTgene_inTissue)
# Define the save parameters
base_save_parameters <- list(
  save_dir = results_folder,  # Directory where the plots will be saved
  save_format = "png",        # Save as PNG format
  units = "in",               # Units for width and height (inches)
  base_width = 10,             # Width of the plot
  base_height = 10,            # Height of the plot
  dpi = 600                   # Resolution (DPI)
)

```


## 30uM7738
```{r}
SpatialTrans_30uM7738_ONTgene_inTissue <- spatPlot2D(gobject = SpatialTrans_30uM7738_ONTgene, 
           cell_color = "in_tissue",
           show_image = TRUE, 
           point_size = 2.5,
           cell_color_code = c("0" = "lightgrey", "1" = "blue"),
           point_alpha = 0.5,
           show_plot = TRUE )

print(SpatialTrans_30uM7738_ONTgene_inTissue)
```

# Join Giotto Objects

```{r}
SpatialTrans_combined_ONTgene <- joinGiottoObjects(gobject_list = list(SpatialTrans_DMSO_ONTgene,SpatialTrans_30uM7738_ONTgene),
                                   gobject_names = c("DMSO", "30uM7738"),
                                   join_method = "shift", x_padding = -10000)

# Printing the file structure for the individual datasets
print(head(pDataDT(SpatialTrans_combined_ONTgene)))
print(SpatialTrans_combined_ONTgene)
```

```{r}
# Create the directory if it doesn't exist
if (!dir.exists(results_folder)) {
  dir.create(results_folder, recursive = TRUE)
}

# Define the save parameters
base_save_parameters <- list(
  save_dir = results_folder,  # Directory where the plots will be saved
  save_format = "png",        # Save as PNG format
  units = "in",               # Units for width and height (inches)
  base_width = 10,            # Width of the plot
  base_height = 10,           # Height of the plot
  dpi = 600                   # Resolution (DPI)
)


SpatialTrans_combined_ONTgene_inTissue <- spatPlot2D(gobject = SpatialTrans_combined_ONTgene, 
           cell_color = "in_tissue", 
           cell_color_code = c("0" = "lightgrey", "1" = "blue"),
           show_image = TRUE, 
           image_name = c("DMSO-image", "30uM7738-image"), 
           point_size = 1, 
           point_alpha = 0.5,
           save_param = base_save_parameters)
print(SpatialTrans_combined_ONTgene_inTissue)
```

12.8 Analyzing joined objects
12.8.1 Normalization and adding statistics

```{r}

# subset on in-tissue spots
metadata <- pDataDT(SpatialTrans_combined_ONTgene )
in_tissue_barcodes <- metadata[in_tissue == 1]$cell_ID
SpatialTrans_combined_ONTgene  <- subsetGiotto(SpatialTrans_combined_ONTgene , 
                              cell_ids = in_tissue_barcodes)

## filter
SpatialTrans_combined_ONTgene  <- filterGiotto(gobject = SpatialTrans_combined_ONTgene ,
                              expression_threshold = 1,
                              feat_det_in_min_cells = 170,
                              min_det_feats_per_cell = 550,
                              expression_values = "raw",
                              verbose = TRUE)

## normalize
SpatialTrans_combined_ONTgene  <- normalizeGiotto(gobject = SpatialTrans_combined_ONTgene , 
                                 scalefactor = 6000)

## add gene & cell statistics
SpatialTrans_combined_ONTgene  <- addStatistics(gobject = SpatialTrans_combined_ONTgene , 
                               expression_values = "raw")

## visualize
SpatialTrans_combined_ONTgene_fil <- spatPlot2D(gobject = SpatialTrans_combined_ONTgene, 
           cell_color = "nr_feats", 
           color_as_factor = FALSE, 
           point_size = 3, 
           show_image = TRUE, 
           image_name = c("DMSO-image","30uM7738-image"),
           show_plot = TRUE,
           save_param = base_save_parameters)
print(SpatialTrans_combined_ONTgene_fil)
```


```{r}

gene_metadata <- fDataDT(SpatialTrans_combined_ONTgene)
testfeats <- gene_metadata[perc_cells > 20 & perc_cells < 50][100:110]$feat_ID

violinPlot(SpatialTrans_combined_ONTgene, 
           feats = testfeats, 
           cluster_column = "list_ID")


```

#Clustering the datasets

```{r}

## PCA ##
SpatialTrans_combined_ONTgene  <- calculateHVF(gobject = SpatialTrans_combined_ONTgene )

SpatialTrans_combined_ONTgene  <- runPCA(gobject = SpatialTrans_combined_ONTgene , 
                        center = TRUE, 
                        scale_unit = TRUE)

## cluster and run UMAP ##
# sNN network (default)
SpatialTrans_combined_ONTgene  <- createNearestNetwork(gobject = SpatialTrans_combined_ONTgene ,
                                      dim_reduction_to_use = "pca", 
                                      dim_reduction_name = "pca",
                                      dimensions_to_use = 1:10, 
                                      k = 15)

# Leiden clustering
SpatialTrans_combined_ONTgene  <- doLeidenCluster(gobject = SpatialTrans_combined_ONTgene , 
                                 resolution = 0.2,
                                 n_iterations = 200)

# UMAP
SpatialTrans_combined_ONTgene  <- runUMAP(SpatialTrans_combined_ONTgene )

```

```{r}
spatDimPlot2D(gobject = SpatialTrans_combined_ONTgene,
              cell_color = "leiden_clus", 
              show_image = TRUE, 
              image_name = c("DMSO-image", "30uM7738-image"),
              save_param = base_save_parameters,
              return_plot =TRUE)

```

# Perform Harmony and default workflows

```{r}
library(harmony)

## run harmony integration
SpatialTrans_combined_ONTgene  <- runGiottoHarmony(SpatialTrans_combined_ONTgene , 
                                  vars_use = "list_ID",
                                  do_pca = FALSE,
                                  sigma = 0.1,
                                  theta = 2,
                                  lambda = 1,
                                  nclust = NULL)

```

# 12.9.1 Clustering harmonized object

```{r}
## sNN network (default)
SpatialTrans_combined_ONTgene  <- createNearestNetwork(gobject = SpatialTrans_combined_ONTgene ,
                                      dim_reduction_to_use = "harmony",
                                      dim_reduction_name = "harmony", 
                                      name = "NN.harmony",
                                      dimensions_to_use = 1:10, 
                                      k = 15)

## Leiden clustering
SpatialTrans_combined_ONTgene  <- doLeidenCluster(gobject = SpatialTrans_combined_ONTgene ,
                                 network_name = "NN.harmony", 
                                 resolution = 0.2, 
                                 n_iterations = 1000, 
                                 name = "leiden_harmony")

# UMAP dimension reduction
SpatialTrans_combined_ONTgene  <- runUMAP(SpatialTrans_combined_ONTgene , 
                         dim_reduction_name = "harmony", 
                         dim_reduction_to_use = "harmony", 
                         name = "umap_harmony")

spatDimPlot2D(gobject = SpatialTrans_combined_ONTgene ,
              dim_reduction_to_use = "umap", 
              dim_reduction_name = "umap_harmony",
              cell_color = "leiden_harmony", 
              #show_image = TRUE, 
              image_name = c("DMSO-image",  "30uM7738-image"),
              spat_point_size = 1, 
              save_param = base_save_parameters)
```

# Vizualizing the tissue contribution

```{r step12_tissue_contribution}

spatDimPlot2D(gobject = SpatialTrans_combined_ONTgene ,
              dim_reduction_to_use = "umap", 
              dim_reduction_name = "umap_harmony",
              cell_color = "list_ID", 
              save_plot = TRUE,
              save_param = base_save_parameters)
```

# Supplementary Figure S6: Harmony Batch Correction Assessment

Assessment of batch effects before and after Harmony integration:

- (A) Pre-Harmony Leiden clustering showing clusters segregated by tissue section
- (B) Post-Harmony Leiden clustering demonstrating improved mixing across conditions
- (C) Post-Harmony tissue contribution confirming balanced representation

```{r figS6_harmony_assessment, fig.width=14, fig.height=8}
library(ggplot2)
library(patchwork)
library(cowplot)

# Generate UMAP plots separately with minimal margins
umap_A <- plotUMAP(gobject = SpatialTrans_combined_ONTgene,
                   dim_reduction_name = "umap",
                   cell_color = "leiden_clus",
                   point_size = 2,
                   title = "",
                   return_plot = TRUE,
                   save_plot = FALSE) + 
  theme(plot.margin = margin(0, 0, 0, 0),
        legend.position = "right",
        legend.key.size = unit(0.3, "cm"))

umap_B <- plotUMAP(gobject = SpatialTrans_combined_ONTgene,
                   dim_reduction_name = "umap_harmony",
                   cell_color = "leiden_harmony",
                   point_size = 2,
                   title = "",
                   return_plot = TRUE,
                   save_plot = FALSE) + 
  theme(plot.margin = margin(0, 0, 0, 0),
        legend.position = "right",
        legend.key.size = unit(0.3, "cm"))

umap_C <- plotUMAP(gobject = SpatialTrans_combined_ONTgene,
                   dim_reduction_name = "umap_harmony",
                   cell_color = "list_ID",
                   point_size = 2,
                   title = "",
                   return_plot = TRUE,
                   save_plot = FALSE) + 
  theme(plot.margin = margin(0, 0, 0, 0),
        legend.position = "right",
        legend.key.size = unit(0.3, "cm"))

# Generate spatial plots separately with minimal margins
spat_A <- spatPlot2D(gobject = SpatialTrans_combined_ONTgene,
                     cell_color = "leiden_clus",
                     show_image = TRUE,
                     image_name = c("DMSO-image", "30uM7738-image"),
                     point_size = 2,
                     title = "",
                     return_plot = TRUE,
                     save_plot = FALSE) + 
  theme(plot.margin = margin(0, 0, 0, 0),
        legend.position = "right",
        legend.key.size = unit(0.3, "cm"))

spat_B <- spatPlot2D(gobject = SpatialTrans_combined_ONTgene,
                     cell_color = "leiden_harmony",
                     show_image = TRUE,
                     image_name = c("DMSO-image", "30uM7738-image"),
                     point_size = 2,
                     title = "",
                     return_plot = TRUE,
                     save_plot = FALSE) + 
  theme(plot.margin = margin(0, 0, 0, 0),
        legend.position = "right",
        legend.key.size = unit(0.3, "cm"))

spat_C <- spatPlot2D(gobject = SpatialTrans_combined_ONTgene,
                     cell_color = "list_ID",
                     show_image = TRUE,
                     image_name = c("DMSO-image", "30uM7738-image"),
                     point_size = 2,
                     title = "",
                     return_plot = TRUE,
                     save_plot = FALSE) + 
  theme(plot.margin = margin(0, 0, 0, 0),
        legend.position = "right",
        legend.key.size = unit(0.3, "cm"))

# Add A, B, C labels directly to UMAP plots
umap_A_labeled <- umap_A + labs(tag = "A") + theme(plot.tag = element_text(size = 14, face = "bold"))
umap_B_labeled <- umap_B + labs(tag = "B") + theme(plot.tag = element_text(size = 14, face = "bold"))
umap_C_labeled <- umap_C + labs(tag = "C") + theme(plot.tag = element_text(size = 14, face = "bold"))

# Combine: UMAP row on top, spatial row on bottom (no separate label row)
fig_S6 <- (umap_A_labeled | umap_B_labeled | umap_C_labeled) /
          (spat_A | spat_B | spat_C) +
  plot_layout(heights = c(1, 1.2)) +
  plot_annotation(
    title = "Supplementary Figure S6: Harmony Batch Correction Assessment",
    theme = theme(
      plot.title = element_text(size = 12, face = "bold", hjust = 0.5),
      plot.margin = margin(2, 2, 2, 2)
    )
  ) &
  theme(plot.margin = margin(1, 1, 1, 1))

print(fig_S6)

# Save Figure S6
ggsave(filename = "./figures/paper_figures/supplementary/figure_S6_harmony_assessment.png",
       plot = fig_S6,
       width = 14, height = 8, dpi = 600)

cat("Figure S6 saved to: ./figures/paper_figures/supplementary/figure_S6_harmony_assessment.png\n")
```


# Interactively Plot regions and save their coordinates
```{r}

polygon_coordinates <- plotInteractivePolygons(SpatialTrans_combined_ONTgene_fil)

giotto_polygons <- createGiottoPolygonsFromDfr(polygon_coordinates, 
                                               name = "selections",
                                               calc_centroids = TRUE)

```

# Add the polygons to the Giotto object

```{r}
SpatialTrans_combined_ONTgene <- addGiottoPolygons(gobject = SpatialTrans_combined_ONTgene,
gpolygons = list(giotto_polygons))

```

# Add the corresponding polygon IDs to the cell metadata

```{r}
SpatialTrans_combined_ONTgene <- addPolygonCells(SpatialTrans_combined_ONTgene,
polygon_name = "selections")
```

Extract the coordinates and IDs from cells located within one or multiple regions of interest.

```{r}
getCellsFromPolygon(SpatialTrans_combined_ONTgene,
                    polygon_name = "selections",
                    polygons = "DMSO")
```

# If no polygon name is provided, the function will retrieve cells located within all polygons
```{r}
getCellsFromPolygon(SpatialTrans_combined_ONTgene,
                    polygon_name = "selections")

```

Compare the expression levels of some genes of interest between the selected regions

```{r}
comparePolygonExpression(SpatialTrans_combined_ONTgene,
                         selected_feats = c("SNHG25"))

```

# PseudoBulk Differential Expression Analysis
Calculate the top genes expressed within each region, then provide the result to compare polygons

```{r}
library(data.table)

scran_results <- findMarkers_one_vs_all(
  SpatialTrans_combined_ONTgene,
  spat_unit = "cell",
  feat_type = "rna",
  method = "scran",
  expression_values = "normalized",
  cluster_column = "list_ID",
  min_feats = 2)

top_genes <- scran_results[, head(.SD, 20), by = "cluster"]$feats
# Define the file path for saving

# Define the new save directory
supplementary_figures <- "./figures/paper_figures/supplementary"
scran_markers_file <- file.path(supplementary_figures, "normilized_PseudoBulk_DE_DMSOvsNUC7738_scran.csv")

# Save scran_markers as a CSV file
fwrite(
  scran_results ,
  file = scran_markers_file
)

View(scran_results )
#comparePolygonExpression(SpatialTrans_combined_ONTgene,
#                         selected_feats = top_genes)
                    

```

```{r}
# Define the save parameters
base_save_parameters <- list(
  save_dir = results_folder,  # Directory where the plots will be saved
  save_format = "png",        # Save as PNG format
  units = "in",               # Units for width and height (inches)
  base_width = 4,             # Width of the plot
  base_height = 20,            # Height of the plot
  dpi = 600                   # Resolution (DPI)
)

```

#
```{r}
# Define the list of top genes
custom_feats <- c(
  "MT-CO2","MT-ND6", "POLR2L", "SNHG25", 
  "HLA-B", "B2M", "SNHG9","GPNMB","CD74", "CSTA"
)
```


# Print the list to confirm
print(custom_feats)

```{r}
violinPlot( SpatialTrans_combined_ONTgene, 
           feats = custom_feats, 
           cluster_column = "list_ID",
           save_param = base_save_parameters)
```

```{r}
plotMetaDataHeatmap(SpatialTrans_combined_ONTgene, 
                    selected_feats = top_genes , 
                    metadata_cols = "list_ID",
                    save_param = modifyList(base_save_parameters, list(base_width = 5, base_height = 20))
                    )

```

# Use the spatPlot arguments to isolate and plot each region.
```{r}
spatPlot2D(SpatialTrans_combined_ONTgene,
           cell_color = "leiden_clus",
           group_by = "list_ID",
           point_size = 2,
           show_legend = FALSE,
           return_plot = TRUE,
           save_param = modifyList(base_save_parameters, list(base_width = 10, base_height = 8))
           )
```


```{r}
# Define the save parameters
base_save_parameters <- list(
  save_dir = results_folder,  # Directory where the plots will be saved
  save_format = "png",        # Save as PNG format
  units = "in",               # Units for width and height (inches)
  base_width = 10,             # Width of the plot
  base_height = 10,            # Height of the plot
  dpi = 600                   # Resolution (DPI)
)

```

```{r}
genes <- c("SNHG25", "HLA-B", "B2M", "MT-CO1", "RPL37","POLR2L")
DMSOvsNUC7738_genes <- spatFeatPlot2D(SpatialTrans_combined_ONTgene,
               expression_values = "scaled",
               #group_by = "list_ID", # Use if comparing/grouping within sample
               feats = genes ,
               point_size = 1,
               point_alpha = 0.7,
               show_image = TRUE,
               cow_n_col = 2,
               gradient_midpoint = 0,
               image_name = c("DMSO-image", "30uM7738-image"),
               save_param = modifyList(base_save_parameters, list(base_width = 10, base_height = 8)),
               default_save_name = "DMSOvsNUC7738_genes")

# Save the combined grid as an image
combined_save_path <- "./figures/paper_figures/supplementary/figure_4_wholesection.png"
ggsave(
  filename = combined_save_path,
  plot = DMSOvsNUC7738_genes,
  width = 10,  # Adjust dimensions as needed
  height = 8, 
  dpi = 600
)

print(DMSOvsNUC7738_genes)

```

# Save the Object and Load Object
```{r}
saveGiotto(SpatialTrans_combined_ONTgene, "./GiottoSuite_combined_ONTgene", overwrite = TRUE)

```

# Load Giotto Object

```{r}
SpatialTrans_combined_ONTgene<- loadGiotto("./GiottoSuite_combined_ONTgene")
```

```{r}
DMSOvsNUC7738_genes <- spatFeatPlot2D(SpatialTrans_combined_ONTgene,
               expression_values = "scaled",
               #group_by = "list_ID", # Use if comparing/grouping within sample
               feats = "SNHG25" ,
               point_size = 1,
               point_alpha = 0.7,
               show_image = TRUE,
               #cow_n_col = 2,
               #image_name = c("DMSO-image", "30uM7738-image"),
               #save_param = modifyList(base_save_parameters, list(base_width = 10, base_height = 8)),
               default_save_name = "DMSOvsNUC7738_genes")
print(DMSOvsNUC7738_genes)
```


```{r}
# Loop through each feature in custom_feats
for (feat in custom_feats) {
  
  # Construct the save name dynamically based on the feature name
  save_name <- paste0("Combo_", feat)
  
  # Run the spatFeatPlot2D function for the current feature
  spatFeatPlot2D(
    SpatialTrans_combined_ONTgene,
    expression_values = "scaled",
    feats = feat,  # Use the current feature
    point_size = 2,
    point_alpha = 0.7,
    show_image = TRUE,
    gradient_midpoint = 0,
    image_name = c("DMSO-image", "30uM7738-image"),
    save_param = base_save_parameters,  # Use the predefined save parameters
    default_save_name = save_name  # Use the dynamically constructed save name
  )
}
```



# Session Information
```{r}
sessionInfo()
```
